---
title: "p8105_hw3_zl3263"
output: github_document
date: "2022-10-12"
---


*import libraries and data*
```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
head(instacart)
```
### Problem 1

The dataset records the order information of instacart website. It has `r nrow(instacart)` rows and `r ncol(instacart)`columns. It has 15 variables containing information like the order id , product id, product names etc. Each row records one product in an order and relevent information both on the order and the product like the date of the order and the department of the product. 

### Problem 2

```{r}
# load and tidy the data
accelerometer = read_csv('data/accel_data.csv')%>%
  janitor::clean_names()%>%
  mutate(isweekend = ifelse(day == c("Saturday", "Sunday"), TRUE, FALSE),.before = activity_1)


```
```{r}
# what does "encode data with reasonable variable classes" mean?
accelerometer_dailytotal = select(accelerometer,colnames(accelerometer)[1:4])
total_activity = rowSums(select(accelerometer,colnames(accelerometer)[5:1444]))
accelerometer_dailytotal$total_activity = total_activity
remove(total_activity) #not elegant but how to optimize?
head(accelerometer_dailytotal)

# tibble with sum generated

```
```{r}
accelerometer_dailytotal %>%
  group_by(day)%>%
  summarize(mean_activity = mean(total_activity))

```
The total activity shows a increasing trend in 7 days from Saturday in previous week to Friday in the next week. 

```{r}
#prepare accelerometer for plotting.
accelerometer_forplot = 
  pivot_longer(
    accelerometer,
    activity_1:activity_1440,
    names_to = "timemark",
    values_to = "activity"
  )%>%
  mutate(timemark = as.integer(substr(timemark,10,13)))

head(accelerometer_forplot)

```
```{r}
accelerometer_forplot$day = factor(accelerometer_forplot$day, 
                                   levels = rev(c('Saturday','Sunday','Monday','Tuesday','Wednesday','Thursday','Friday')))

#why single panel！single panel cannot see anything
ggplot(accelerometer_forplot,aes(x = timemark/60, y = log(activity), color = day)) +
  geom_point(alpha = 0.3, size = 0.5) 

# no problem with this pic but why the data looks like that qaq
  

  
```

